---
title:  "Preparing a WHAM input"
subtitle: ""
author: "Tim Miller<br>NOAA Fisheries, NEFSC<br> 2024-06-03"
output:
  xaringan::moon_reader:
    self_contained: true
    css: ["xaringan-themer_16_9.css", "slides-style_TMB201.css"]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

layout: true

.footnote[U.S. Department of Commerce | National Oceanic and Atmospheric Administration | National Marine Fisheries Service]


<style type="text/css">

code.cpp{
  font-size: 14px;
}
code.r{
  font-size: 14px;
}


</style>

```{css, echo=FALSE}
pre {
  max-height: 250px; /*changes height of chunk output box*/
  max-width: 800px; /*changes width of chunk output box*/
  overflow-y: auto; /* auto will add vertical scroll bar when necessary */
}

```


```{r setup, include=FALSE}
#options(htmltools.dir.version = FALSE)
#library(knitr)
#knitr::opts_chunk$set(cache = TRUE)
#knitr::opts_knit$set(root.dir = here::here())
#knitr::opts_chunk$set(width = 1500)
#library(kableExtra)
#library(imager)
#here::i_am("c:/work/BSB_RT")

#library(here)

```

```{r set-options, include = FALSE}
options(width = 50)
```

```{r xaringan-tile-view, echo=FALSE}
# this gives you a tile navigation if you type "O" at any time
#xaringanExtra::use_tile_view()
```

---
# Outline <br>

Creating an input for fitting a model with WHAM:
* Using ASAP input
* Without ASAP input
* Modifying an existing WHAM input
* Simple multiple-stock and regions
---

# Using ASAP Input <br>

For illustration, we will use one of the ASAP3 dat files that comes with the WHAM package. 

First, load the WHAM package. 
This will find and load the wham package that is first in the search list of locations that R checks.
```{r, eval = FALSE}
library(wham)
```
```{r, include = FALSE, eval = TRUE}
library("wham", lib.loc = "c:/work/wham/old_packages/lab")
```

When there are multiple versions of WHAM installed a particular version can be loaded. 

For example, this will load a version of wham that is installed to the directory `c:/work/wham/old_packages/76093e0`
```{r, eval = FALSE}
library("wham", lib.loc = "c:/work/wham/old_packages/76093e0")
```

---

# Using ASAP Input <br>

Second, locate an ASAP3 dat file available in  the WHAM package and load the information from ASAP3 dat files into R using the `read_asap3_dat` function.

```{r, eval = TRUE}
path_to_examples <- system.file("extdata", package="wham")
asap3 <- read_asap3_dat(file.path(path_to_examples,"ex2_SNEMAYT.dat"))
```
---

# Using ASAP Input <br>

Third, create a basic WHAM input using the `prepare_wham_input` function. 
```{r, eval = TRUE}
input <- prepare_wham_input(asap3) 
```

---

# Using ASAP Input <br>

Third, create a basic WHAM input using the `prepare_wham_input` function. 
```{r, eval = FALSE}
input <- prepare_wham_input(asap3) 
```
The input description can be printed at any time by
```{r, eval = FALSE}
cat(unlist(input$log))
```

With no other arguments specified to the function:
* a basic statistical catch-at-age model will be configured. 
 * No random effects
 * Annual recruitment treated as fixed effects.
 * Stock-recruit function is not allowed.
* It will be specified to match the ASAP3 configuration as closely as possible.
 * Same data, CV, effective sample size assumptions.
 * Typically same assumptions for initial parameter values.
 * BUT any ASAP3 penalties are ignored.

---


# Without ASAP Input

First, we need to define all of the necessary data inputs for WHAM. We will use the same information as that in the `ex2_SENMAYT.dat` file and show that the results are the same whether we use the asap object directly or cobble the input together by hand.

First the biological information:

```{r, eval = TRUE}
years <- as.integer(input$years)
ages <- 1:input$data$n_ages
waa <- input$data$waa
mat <- input$data$mature
Maa <- exp(matrix(c(input$par$Mpars), length(years), length(ages), byrow = TRUE) + c(input$par$M_re[1,1,,]))
waa_pointer_indices <- input$data$waa_pointer_indices
waa_pointer_fleets <- input$data$waa_pointer_fleets
waa_pointer_ssb <- input$data$waa_pointer_ssb
fracyr_ssb <-  input$data$fracyr_SSB
```
---

# Without ASAP Input

Next the catch information:
```{r, eval = TRUE}
catch <- input$data$agg_catch
catch_cv <- cbind(sqrt(exp(input$data$agg_catch_sigma^2) - 1))
catch_Neff <- cbind(input$data$catch_Neff)
catch_paa <- input$data$catch_paa
use_catch_paa <- input$data$use_catch_paa
selblock_pointer_fleets <- input$data$selblock_pointer_fleets
```

Finally, the information for all indices:
```{r, eval = TRUE}
index <- input$data$agg_indices
index_cv <- cbind(sqrt(exp(input$data$agg_index_sigma^2) - 1))
index_Neff <- cbind(input$data$index_Neff)
index_fracyr <- input$data$fracyr_indices
units_indices <- input$data$units_indices
units_index_paa <- input$data$units_index_paa
use_indices <- input$data$use_indices
use_index_paa <- input$data$use_index_paa
index_paa <- input$data$index_paa
selblock_pointer_indices <- input$data$selblock_pointer_indices
```

---


# Without ASAP Input

Now we are ready to build the necessary arguments required by `prepapre_wham_input`. 

First is the `basic_info` argument which includes biological information. Note, we now provide information on the number of stocks, regions, and intervals, but default values are the same.
```{r, eval = TRUE}
basic_info <- list(
  n_stocks = 1L,
  ages = ages,
  n_seasons = 1L,
  n_fleets = 1L,
  fracyr_SSB = fracyr_ssb,
  maturity = mat,
  years = years,
  waa = waa,
  waa_pointer_ssb = waa_pointer_ssb
)

```

---

# Without ASAP Input

Second, the catch information:
```{r, eval = TRUE}
catch_info <- list(
  n_fleets = NCOL(catch),
  agg_catch = catch,
  agg_catch_cv = catch_cv,
  catch_paa = catch_paa,
  use_catch_paa = use_catch_paa,
  catch_Neff = catch_Neff,
  selblock_pointer_fleets = selblock_pointer_fleets,
  waa_pointer_fleets = waa_pointer_fleets
)

```
---

# Without ASAP Input

Next, the information for the indices:
```{r, eval = TRUE}
index_info <- list(
  n_indices = NCOL(index),
  agg_indices = index,
  units_indices = units_indices,
  units_index_paa = units_index_paa,
  agg_index_cv = index_cv,
  fracyr_indices = index_fracyr,
  use_indices = use_indices,
  use_index_paa = use_index_paa,
  index_paa = index_paa,
  index_Neff = index_Neff,
  selblock_pointer_indices = selblock_pointer_indices,
  waa_pointer_indices = waa_pointer_indices
)
```
---

# Without ASAP Input

The last two arguments to prepare_wham_input that we need are for natural mortality and selectivity. Natural mortality is separated from basic biology because there are many different ways to model natural mortality in WHAM.

Natural mortality is not estimated. Assumed M changes with age for this stock, but is constant over time.
```{r, eval = TRUE}
M_in <- list(
  initial_MAA = array(Maa, dim = c(1,1,length(basic_info$years), length(basic_info$ages)))
)
```

Specify using logistic selectivity for all 6 selectivity blocks and initial values for a50 and inverse of slope. Fixing selectivity for the last two selectivity blocks.
```{r, eval = TRUE}
selectivity <- list(
  model = rep("logistic",6), n_selblocks = 6,
  fix_pars = list(NULL,NULL,NULL,NULL, 1:2, 1:2),
  initial_pars = list(c(2,0.2),c(2,0.2),c(2,0.2),c(2,0.2),c(1.5,0.1),c(1.5,0.1))
)
```

---

# Creating WHAM inputs: 3 ways

All at once
```{r, eval = FALSE}
input_all <- prepare_wham_input(basic_info = basic_info, selectivity = selectivity, catch_info = catch_info, index_info = index_info, M = M_in)
```

Build input sequentially
```{r, eval = FALSE}
input_seq <- prepare_wham_input(basic_info = basic_info)
input_seq <- set_M(input_seq, M = M_in)
input_seq <- set_catch(input_seq, catch_info = catch_info)
input_seq <- set_F(input_seq)
input_seq <- set_indices(input_seq, index_info = index_info)
input_seq <- set_q(input_seq)
input_seq <- set_selectivity(input_seq, selectivity = selectivity)
```
Using asap input
```{r, eval = FALSE}
input_asap <- prepare_wham_input(asap3) 
```
```{r, include = FALSE}
input_all <- prepare_wham_input(basic_info = basic_info, selectivity = selectivity, catch_info = catch_info, index_info = index_info, M = M_in)
input_seq <- prepare_wham_input(basic_info = basic_info)
input_seq <- set_M(input_seq, M = M_in)
input_seq <- set_catch(input_seq, catch_info = catch_info)
input_seq <- set_F(input_seq)
input_seq <- set_indices(input_seq, index_info = index_info)
input_seq <- set_q(input_seq)
input_seq <- set_selectivity(input_seq, selectivity = selectivity)
input_asap <- prepare_wham_input(asap3) 
```
---

# Inside the WHAM input

The object returned by `prepare_wham_input` is a list of objects. 
```{r, eval = TRUE}
names(input_asap)
```
* Some elements are used by `TMB::MakeADFun`: data, par, map, random
* Others provide information for generating output: years, years_full, ages.lab, index_names, fleet_names, etc. 
* `log` is the list of text strings printed to the screen by `prepare_wham_input`
* `options` are the lists provided for the various arguments to `prepare_wham_input` or the various `set_*` functions.
* `call` saves the call expression to `prepare_wham_input`

---

# Compare unfit models


```{r, eval = FALSE}
nofit_all <- fit_wham(input_all, do.fit = FALSE)
nofit_seq <- fit_wham(input_seq, do.fit = FALSE)
nofit_asap <- fit_wham(input_asap, do.fit = FALSE)
```

```{r, include = FALSE}
nofit_all <- fit_wham(input_all, do.fit = FALSE)
nofit_seq <- fit_wham(input_seq, do.fit = FALSE)
nofit_asap <- fit_wham(input_asap, do.fit = FALSE)
```

We can see that the negative log likelihood is the same whether we build the input all at once or sequentially. 
```{r, eval = TRUE}
nofit_all$fn() - nofit_seq$fn() #equal
nofit_all$fn() - nofit_asap$fn() #different
```
---

# Compare unfit models

But building from the asap file provides a different value.

The asap file includes different initial values for initial abuandance at age and catchability:
```{r, eval = TRUE}
nofit_asap$par-nofit_all$par
```
Evaluating from the same initial parameters give the same nll:
```{r, eval = TRUE}
nofit_asap$fn(nofit_all$par)-nofit_all$fn()
```

---

# Compare fitted models

And you can verify that the optimized models using each of the 3 inputs gives the same results.

```{r, eval = FALSE}
fit_seq <- fit_wham(input_seq, do.retro = FALSE, do.osa = FALSE, do.sdrep = FALSE)
fit_all <- fit_wham(input_all, do.retro = FALSE, do.osa = FALSE, do.sdrep = FALSE)
fit_asap <- fit_wham(input_asap, do.retro = FALSE, do.osa = FALSE, do.sdrep = FALSE)
```
```{r, include = FALSE}
res_dir <- file.path("../temp")
fit_seq <- readRDS(file.path(res_dir,"fit_seq.RDS"))
fit_all <- readRDS(file.path(res_dir,"fit_all.RDS"))
fit_asap <- readRDS(file.path(res_dir,"fit_asap.RDS"))
```

```{r, eval = TRUE}
fit_seq$opt$obj - fit_all$opt$obj # equal
fit_asap$opt$obj - fit_all$opt$obj # equal
```

---

#Multiple stocks

A contrived example: Estimating the same model two times simultaneously.
```{r, eval = FALSE}
two_stocks_asap <- read_asap3_dat(file.path(path_to_examples,c("ex1_SNEMAYT.dat","ex1_SNEMAYT.dat")))
ini_2_stocks <- prepare_wham_input(two_stocks_asap)
fit_2_stocks <- fit_wham(ini_2_stocks, do.osa = FALSE, do.retro= FALSE, do.sdrep = FALSE)
```
```{r, include = FALSE}
fit_2_stocks_RDS <- file.path(res_dir,"fit_2_stocks.RDS")
fit_2_stocks <- readRDS(fit_2_stocks_RDS)
```
Model estimates are identical for the two stocks, as they should be.
```{r, eval = TRUE, max.height='10px'}
fit_2_stocks$rep$SSB
```
