---
title:  "Reference points and Projections"
subtitle: ""
author: "Tim Miller<br>NOAA Fisheries, NEFSC<br> 2024-06-4"
output:
  xaringan::moon_reader:
    self_contained: true
    css: ["xaringan-themer_16_9.css", "slides-style_TMB201.css"]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

layout: true

.footnote[U.S. Department of Commerce | National Oceanic and Atmospheric Administration | National Marine Fisheries Service]


<style type="text/css">

code.cpp{
  font-size: 14px;
}
code.r{
  font-size: 14px;
}


</style>

```{css, echo=FALSE}
pre {
  max-height: 250px; /*changes height of chunk output box*/
  max-width: 800px; /*changes width of chunk output box*/
  overflow-y: auto; /* auto will add vertical scroll bar when necessary */
}

```
```{r set-options, include = FALSE}
options(width = 50)
```

```{r xaringan-tile-view, echo=FALSE}
# this gives you a tile navigation if you type "O" at any time
#xaringanExtra::use_tile_view()
```

---

# Outline <br>

* Using `basic_info` argument to `prepare_wham_input`
* Using different reference point configuration tools
* Using `project_wham`
* Using different projection options
---

# Abundance at age options
The `NAA_re` argument to `prepare_wham_input` or `set_NAA` has these elements to configure selectivity.
* `recruit_model`: whether to estimate recruitment as
 * (1) annual fixed effects
 * (2) a mean recruitment parameter and annual random effects,
 * (3) a mean Beverton-Holt function of SSB and annual random effects
 * (4) a mean Ricker function of SSB and annual random effects
* `recruit_pars`: initial values to use for mean recruitment parameters.
* `N1_model`: character vector specifying which of 4 options for the initial abundance at age
 * `age-specific`: a different parameter for each age class
 * `equilibrium`: 2 parameters: initial recruitment and a fully-selected F needed to calculate an equilibrium distribution of abundance at age.
 * `iid-re`: Estimate a mean abundance with independent random effects for each age.
 * `ar1-re`: Estimate a mean abundance with autocrrelated random effects for each age.
* `N1_pars`: initial parameter values to use for the initial abundance at age.

---

# Abundance at age options
* `sigma`: whether to estimate random effects on recruitment only or recruitment and survival from one time step to the next.
* `recruit_pars`: initial values to use for mean recruitment parameters.
* `N1_model`: character vector specifying which of 4 options for the initial abundance at age
 * `age-specific`: a different parameter for each age class
 * `equilibrium`: 2 parameters: initial recruitment and a fully-selected F needed to calculate an equilibrium distribution of abundance at age.
 * `iid-re`: Estimate a mean abundance with independent random effects for each age.
 * `ar1-re`: Estimate a mean abundance with autocrrelated random effects for each age.
* `N1_pars`: initial parameter values to use for the initial abundance at age.

---
