---
title:  "Review and Next Steps"
subtitle: ""
author: "Tim Miller<br>NOAA Fisheries, NEFSC<br> 2024-06-07"
output:
  xaringan::moon_reader:
    self_contained: true
    css: ["xaringan-themer_16_9.css", "slides-style_TMB201.css"]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

layout: true

.footnote[U.S. Department of Commerce | National Oceanic and Atmospheric Administration | National Marine Fisheries Service]


<style type="text/css">

code.cpp{
  font-size: 14px;
}
code.r{
  font-size: 14px;
}


</style>

```{css, echo=FALSE}
pre {
  max-height: 250px; /*changes height of chunk output box*/
  max-width: 800px; /*changes width of chunk output box*/
  overflow-y: auto; /* auto will add vertical scroll bar when necessary */
}

```
```{r set-options, include = FALSE}
options(width = 50)
```

```{r xaringan-tile-view, echo=FALSE}
# this gives you a tile navigation if you type "O" at any time
#xaringanExtra::use_tile_view()
```

---

# Outline <br>

* Review what we have covered this week
* Did not cover 
* How everyone can contribute!

---

# Getting data into WHAM
---

# Configuring recruitment and abundance at age
---

# Configuring selectivity
---

# Configuring natural mortality
---

# Configuring catchability
---

# Reference points
---

# Projections
---

# What we did not cover

* different age composition likelihood options
* multi-stock, multi-area features

---

# Contributing to help improve WHAM

* First, you need to have 
    * git installed
    * a github account
* You can use git and github through Rstudio or many other coding environments
* You can help without knowing C++ and TMB!
* Developing on the R side can help with
    * Summarizing input and output from models
    * Summarizing model configuration specified by the user
    * Generating output to facilitate reporting for management

---

# Steps to contributing

* First clone the repo as a `wham` directory. 
* Using the command line from an appropriate parent directory:
    * `git clone https://github.com/timjmiller/wham.git`
* Now you have the WHAM repo!
* Switch to devel branch
    * `git checkout devel`
* Create a branch off of devel appropriately named (e.g., "new_branch_name") for what you want to add
    * `git branch new_banch_name`
* Then switch to your new branch
    *  `git checkout new_branch_name`
* Make all amazing changes to WHAM
* Make useful tests of these changes for future development
* When devtools::test() completes successfully submit pull request!
* If you intend only to add a useful R function, just need to compile the model 
    * From an R session in the wham directory run `pkgbuild::compile_dll(debug = FALSE)`
    * Then `pkgload::load_all()` will load the WHAM package from your local copy of the repo
    * During development of R functions can repeatedly call `pkgload::load_all()` to update your package
    * See Roxygen code in other WHAM functions to see how to develop useful help files for the new function(s).