---
title:  "Review and Going Forward"
subtitle: ""
author: "Tim Miller<br>NOAA Fisheries, NEFSC<br> 2024-06-07"
output:
  xaringan::moon_reader:
    self_contained: true
    css: ["xaringan-themer_16_9.css", "slides-style_TMB201.css"]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

layout: true

.footnote[U.S. Department of Commerce | National Oceanic and Atmospheric Administration | National Marine Fisheries Service]


<style type="text/css">

code.cpp{
  font-size: 14px;
}
code.r{
  font-size: 14px;
}


</style>

```{css, echo=FALSE}
pre {
  max-height: 250px; /*changes height of chunk output box*/
  max-width: 800px; /*changes width of chunk output box*/
  overflow-y: auto; /* auto will add vertical scroll bar when necessary */
}

```
```{r set-options, include = FALSE}
options(width = 50)
```

```{r xaringan-tile-view, echo=FALSE}
# this gives you a tile navigation if you type "O" at any time
#xaringanExtra::use_tile_view()
```

---

# Outline <br>

* Review what we have covered this week
* Did not cover 
* How everyone can contribute!

---

# Getting data into WHAM

* Most people are using ASAP files
* Learned how to create input without ASAP files
    * Good for future without ASAP
* Can adjust existing inputs using `set_*` functions

# Configuring selectivity

* Mean selectivity models
* Random effects are on the mean model parameters (not necessarily on ages)
* Can be used to model FAA varying over time like SAM

---

# Configuring recruitment and abundance at age

* More options for initial numbers at age now
* Configuring NAA random effects pretty much the same as before
    * Except no lognormal adjustment and recruitment is decoupled by default


# Configuring natural mortality

* Options for estimating constant natural mortality
  * by Age?
  * For some ages and not others?
  * Estimate M for ages 1-3 but make it the same for all ages?

# Configuring catchability

* time-varying catchability might be helpful for
    * indices that cover a small portion of the stock area
    * species with life styles that are not observed well by trawl surveys

---

# Debugging WHAM

* Use nll reporting to deduce where NA issues are.
* Re-examine CVs and effective sample sizes from ASAP

# Environmental effects

* If observation error estimates available use them
    * estimating observation SD can provide rosy picture of effects of covariates
* Should compare models with and without covariates that both have residual variability
    * Happens automatically with recruitment, but not as obvious for other things

# Reference points and projections

* Need to be careful of XSPR_R_opt for prevailing static reference points
* Lots of options for configuring projections for 
    * F, Catch, M, Covariates, WAA, maturity

---

# Simulation

* Endless possibilities
    * Self tests
    * Cross tests
    * Simulating an imaginary fish stock
    * Closed-loop simulation

# What we did not cover

* multi-stock, multi-area features
* details of reference points and projections with multiple fleets

---

# Contributing to help improve WHAM

* Needs 
    * git installed
    * a github account
* You can use git and github through Rstudio or many other coding environments
* You can help without knowing C++ and TMB!
* Developing on the R side can help with
    * Summarizing input and output from models
    * Summarizing model configuration specified by the user
    * Generating output to facilitate reporting for management

---

# Steps to contributing

* First clone the repo as a `wham` directory. 
    * Using the command line from an appropriate parent directory:
        * `git clone https://github.com/timjmiller/wham.git`
    * Now you have the WHAM repo!
* Switch to devel branch (or other branch to work from)
    * `git checkout devel`
* Create a branch off of devel appropriately named (e.g., "new_branch_name") for what you want to add
    * `git branch new_banch_name`
* Then switch to your new branch
    *  `git checkout new_branch_name`
* Make all amazing changes to WHAM
    * E.g., write a new R function that makes a table with columns: Year, SSB, R, F and corresponding SD, and CIs  


---

# Steps to contributing


* Make some simple tests of these changes for future development
    * see `tests` directory of repo for inspiration
* Make sure your branch is up to date with parent branch
    * `git fetch`
    * `git merge devel` (resolve any merge conflicts)
* Submit pull request when
    * `devtools::test()` completes successfully 
    * `devtools::document()` completes successfully
* Workflow: 
    * If you intend only to add a useful R function, just need to compile the model once.
        * From an R session in the wham directory run `pkgbuild::compile_dll(debug = FALSE)`
        * Then `pkgload::load_all()` will load the WHAM package from your local copy of the repo
        * During development of R function(s) can repeatedly call `pkgload::load_all()` to update your package with changes to the R function(s).
        * See Roxygen code in other WHAM functions to see how to write help files for the new function(s).
    * Developing TMB features is an advanced level of contribution, 
        * Need to successfully compile the model with the changes.
        * Overcoming C++ bugs causing failed compilation is advanced level contribution. 
