---
title:  "Simulation"
subtitle: ""
author: "Tim Miller<br>NOAA Fisheries, NEFSC<br> 2024-06-07"
output:
  xaringan::moon_reader:
    self_contained: true
    css: ["xaringan-themer_16_9.css", "slides-style_TMB201.css"]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

layout: true

.footnote[U.S. Department of Commerce | National Oceanic and Atmospheric Administration | National Marine Fisheries Service]


<style type="text/css">

code.cpp{
  font-size: 14px;
}
code.r{
  font-size: 14px;
}


</style>

```{css, echo=FALSE}
pre {
  max-height: 250px; /*changes height of chunk output box*/
  max-width: 800px; /*changes width of chunk output box*/
  overflow-y: auto; /* auto will add vertical scroll bar when necessary */
}

```
```{r set-options, include = FALSE}
options(width = 50)
```

```{r xaringan-tile-view, echo=FALSE}
# this gives you a tile navigation if you type "O" at any time
#xaringanExtra::use_tile_view()
```

---

# Outline <br>

* Simulate data from a fitted model with input derived from an ASAP3 dat file.
* Make an operating model, simulate data, and fit models.
* Fit models to data simulated from a different model.
* Configure different assumptions about the population for both the operating model and the estimating model.
* Do a closed-loop simulation with operating model, fitting an estimating model, generating catch advice and incorporating it into the operating model.
* **See the simulation vignette for this wham version for greater details**
---

# Getting started

Load some useful packages and use an asap3 dat file from WHAM 
```{r, include = FALSE}
library("wham", lib.loc = "c:/work/wham/old_packages/lab")
wham.dir <- find.package("wham")
asap3 <- read_asap3_dat(file.path(wham.dir,"extdata","ex1_SNEMAYT.dat"))
```
```{r, eval = FALSE}
library(wham)
library(ggplot2)
library(tidyr)
library(dplyr)
wham.dir <- find.package("wham")
asap3 <- read_asap3_dat(file.path(wham.dir,"extdata","ex1_SNEMAYT.dat"))
```

---

# Fit the model

Make an input and fit a model in wham.
```{r, include = FALSE}
selectivity <- list(model = rep("logistic", 3))
selectivity$initial_pars <- list(
  c(3, 0.2), 
  c(3, 0.2), 
  c(3, 0.2)) 

NAA_re = list(
  sigma="rec", #random effects for recruitment only
  cor="iid", #random effects are independent
  recruit_model = 2) #mean recruitment is the only fixed effect parameter
```
```{r, eval=FALSE}
#define selectivity model
selectivity <- list(model = rep("logistic", 3))
selectivity$initial_pars <- list(
  c(3, 0.2), 
  c(3, 0.2), 
  c(3, 0.2)) 

NAA_re = list(
  sigma="rec", #random effects for recruitment only
  cor="iid", #random effects are independent
  recruit_model = 2) #mean recruitment is the only fixed effect parameter

input <- prepare_wham_input(
  asap3, 
  selectivity = selectivity,
  NAA_re = NAA_re) 

fit_1 <- fit_wham(input, do.sdrep = FALSE, do.osa = FALSE, do.retro=FALSE) # don't do retro peels and OSA residuals, don't show output during optimization
```

---
# TMB simulate function

The important item here is `fit_1$simulate`
```{r, include=FALSE}
fit_1 <- readRDS(file.path("../temp", "day_5_1_fit_1.RDS"))
fit_1$simulate
```

```{r, eval=TRUE}
fit_1$simulate
```

---

# TMB simulate function

```{r, eval=FALSE}
fit_1$simulate
```

* This is the workhorse function used by the `self_test` function we explored previously.
* Running this function will 
 * report anything using the `REPORT` calls
 * most importantly, those within SIMULATE statements in the compiled cpp code
* Using the `complete = TRUE` argument tells the function to provide all data inputs whether they are simulated or not.
 * useful for forming inputs for estimation models.

---

# A simple function to simulate and fit models

The below function will take an operating model (om) and an input that may not match the om
* BUT needs the same number of years, fleets, indices, ages
* All observations simulated from the `om` are placed in the `input`
* `obsvec` MUST be included because it is what is used to evaluate the log-likelihood
* the input or fitted estimating model (em) may be returned
* *BOTH observations AND random effects (process errors) are simulated* 

```{r, eval=TRUE}
sim_fn = function(om, input, do.fit = FALSE){
  obs_names = c("agg_indices","agg_catch","catch_paa","index_paa", "Ecov_obs", "obsvec")
  om_data = om$simulate(complete = TRUE)
  input$data[obs_names] = om_data[obs_names]
  input <- set_osa_obs(input) #to transform catch and index paa appropriately
  if(do.fit) {
    fit = fit_wham(input, do.sdrep = FALSE, do.osa = FALSE, do.retro = FALSE, MakeADFun.silent = TRUE)
    return(list(fit, om_data))
  } else return(list(input,om_data))
}
```
---

# An em with an alternative age composition likelihood

* The om assumes (implicitly) that the age composition observations have a multinomial distribution
* Here is an input that instead assumes a logistic normal (treating 0s as missing).

```{r, eval=TRUE, results = "hide"}
input_em <- prepare_wham_input(
  asap3, 
  selectivity=selectivity,
  NAA_re = NAA_re,
  age_comp = "logistic-normal-miss0") #the only difference from original input
```
---

# An em with an alternative age composition likelihood

Using `do.fit = FALSE`, we can see the difference between the age composition observations that are fitted and those that were simulated. 

The 4-9 elements correspond to the index age composition observations in the first year:
```{r, eval = TRUE}
set.seed(123)
sim = sim_fn(fit_1, input_em, do.fit = FALSE)
sim[[1]]$data$obs[1:10,] #data.frame of observations and identifying information
sim[[1]]$data$obsvec[4:9] #showing obsvec that is used (MVN = logistic normal)
```

---

# An em with an alternative age composition likelihood

```{r, eval = TRUE}
sim[[2]]$obsvec[4:9] #showing obsvec that is simulated (multinomial frequencies)
paa <- sim[[2]]$index_paa[1,1,]
paa #multinomial proportions : naa/sum(naa)
sim[[2]]$index_Neff[1] * paa # = simulated obsvec (multinomial frequencies)
log(paa[-6]) - log(paa[6]) # = inverse additive logistic transform for obsvec that is used
sim[[1]]$data$obsvec[4:9] #same as transform of proportions
```

* The multinomial simulates the observations as frequencies. 
* The logistic normal expects proportions, but there is an equivalent multivariate normal transformation that WHAM uses to facilitate OSA residuals. 
* Note that the last age class is NA because the dimension of the MVN is one less than the number of age classes

---

# An em with an alternative age composition likelihood

```{r, eval = FALSE}
set.seed(123)
simfit <- sim_fn(fit_1, input_em, do.fit = TRUE)
```
```{r, include = FALSE}
simfit <- readRDS(file.path("../temp", "day_5_1_simfit_1.RDS"))
```
```{r, eval = TRUE, out.width="40%", fig.align="center"}
plot(fit_1$years, simfit[[2]]$SSB, type = 'l', ylab = "SSB", xlab = "Year")
lines(fit_1$years, simfit[[1]]$rep$SSB, lty = 2, col = 'red')
```

---

# An em with an alternative selectivity model

.pull-left[
* Assume logistic selectivity rather than age-specific

```{r, eval = FALSE}
selectivity_em = list(
  model = c(rep("logistic", input_em$data$n_fleets),rep("logistic", input_em$data$n_indices)),
  initial_pars = rep(list(c(5,1)), input_em$data$n_fleets + input_em$data$n_indices)) #fleet, index

input_em_2 <- prepare_wham_input(
    asap3, 
    selectivity = selectivity_em,
    NAA_re = NAA_re,
    model_name="Ex 1: SNEMA Yellowtail Flounder") 

set.seed(123)
simfit_2 = sim_fn(fit_1, input_em_2, do.fit = TRUE)
plot(fit_1$years, simfit_2[[2]]$SSB, type = 'l', ylab = "SSB", xlab = "Year")
lines(fit_1$years, simfit_2[[1]]$rep$SSB, lty = 2, col = 'red')
```
```{r, include = FALSE}
simfit_2 <- readRDS(file.path("../temp", "day_5_1_simfit_2.RDS"))
```
]
.pull-right[
```{r, echo = FALSE, out.width="100%", fig.align="center"}
plot(fit_1$years, simfit_2[[2]]$SSB, type = 'l', ylab = "SSB", xlab = "Year")
lines(fit_1$years, simfit_2[[1]]$rep$SSB, lty = 2, col = 'red')
```
]
---

# An operating model without an ASAP3 file

* On the first day, we already showed how to generate a WHAM input without an asap file.
* This function does those steps, but also defines a fishing history and a number of base and feedback years for a closed-loop simulation
* It also does not need any real observations because we only need it to simulate a population and data.

```{r, eval=TRUE}
make_info <- function(base_years = 1982:2021, ages = 1:10, Fhist = "updown", n_feedback_years = 0) { #changed years
    info <- list()
    info$ages <- ages
    info$years <- as.integer(base_years[1] - 1 + 1:(length(base_years) + n_feedback_years))
    na <- length(info$ages)
    ny <- length(info$years)
    info$n_regions <- 1L
    info$n_stocks <- 1L
    nby <- length(base_years)
    mid <- floor(nby/2)
    #up then down

    catch_info <- list()
    catch_info$n_fleets <- 1L
    catch_info$catch_cv <- matrix(0.1, ny, catch_info$n_fleets)
    catch_info$catch_Neff <- matrix(200, ny, catch_info$n_fleets)

    F_info <- list()
    if(Fhist == "updown") F_info$F <- matrix(0.2 + c(seq(0,0.4,length.out = mid),seq(0.4,0,length.out=nby-mid)),nby, catch_info$n_fleets)

    #down then up
    if(Fhist == "downup") F_info$F <- matrix(0.2 + c(seq(0.4,0,length.out = mid),seq(0,0.4,length.out=nby-mid)),nby, catch_info$n_fleets)

    if(n_feedback_years>0) F_info$F <- rbind(F_info$F, F_info$F[rep(nby, n_feedback_years),, drop = F]) #same F as terminal year for feedback period

    index_info <- list()
    index_info$n_indices <- 1L
    index_info$index_cv <- matrix(0.3, ny, index_info$n_indices)
    index_info$index_Neff <- matrix(100, ny, index_info$n_indices)
    index_info$fracyr_indices <- matrix(0.5, ny, index_info$n_indices)
    index_info$index_units <- rep(1, length(index_info$n_indices)) #biomass
    index_info$index_paa_units <- rep(2, length(index_info$n_indices)) #abundance
    index_info$q <- rep(0.3, index_info$n_indices)
    
    info$maturity <- array(t(matrix(1/(1 + exp(-1*(1:na - na/2))), na, ny)), dim = c(1, ny, na))

    L <- 100*(1-exp(-0.3*(1:na - 0)))
    W <- exp(-11)*L^3
    nwaa <- index_info$n_indices + catch_info$n_fleets + 2
    info$waa <- array(NA, dim = c(nwaa, ny, na))
    for(i in 1:nwaa) info$waa[i,,] <- t(matrix(W, na, ny))

    info$fracyr_SSB <- cbind(rep(0.25,ny))

    catch_info$selblock_pointer_fleets <- t(matrix(1:catch_info$n_fleets, catch_info$n_fleets, ny))
    index_info$selblock_pointer_indices <- t(matrix(catch_info$n_fleets + 1:index_info$n_indices, index_info$n_indices, ny))
    return(list(basic_info = info, catch_info = catch_info, index_info = index_info, F = F_info))
}
```

---

# An operating model without an ASAP3 file

* We obtain some of the arguments to `prepare_wham_input` and `set_*` functions from the `make_info` function.

```{r, eval=TRUE}
stock_om_info <- make_info()
basic_info <- stock_om_info$basic_info
catch_info <- stock_om_info$catch_info
index_info <- stock_om_info$index_info
F_info <- stock_om_info$F
```

---

# An operating model without an ASAP3 file

* We also need to set up the other arguments to `set_*` functions with initial values that are used to simulate from the operating model
 * logistic selectivity parameters
 * M
 * Initial abundance at age
 * mean recruitment
 * NAA process error parameters
* The ability to set initial values using the arguments to set_* is particularly important for constructing an oM

```{r, eval=TRUE}
selectivity_om <- list(
  model = c(rep("logistic", catch_info$n_fleets),rep("logistic", index_info$n_indices)),
  initial_pars = rep(list(c(5,1)), catch_info$n_fleets + index_info$n_indices)
) #fleet, index

M_om <- list(initial_means = array(0.2, dim = c(1,1, length(basic_info$ages))))

NAA_re_om <- list(
  N1_pars = array(exp(10)*exp(-(0:(length(basic_info$ages)-1))*M_om$initial_means[1]), dim = c(1,1,length(basic_info$ages))),
  sigma = "rec", #random about mean
  cor="iid", #random effects are independent
  recruit_model = 2, #random effects with a constant mean
  recruit_pars = list(exp(10)),
  sigma_vals = array(0.5, c(1,1,length(basic_info$ages))) #sigma_R, only the value in the first age class is used.
)
```

---

# An operating model without an ASAP3 file

* Create the operating model input

```{r, eval=TRUE, results = "hide"}
stock_om_input <- prepare_wham_input(
  basic_info = basic_info, 
  selectivity = selectivity_om, 
  NAA_re = NAA_re_om, 
  catch_info = catch_info,
  index_info = index_info,
  M = M_om,
  F = F_info
)
```
```{r, eval = FALSE}
stock_om = fit_wham(stock_om_input, do.fit = FALSE, MakeADFun.silent = TRUE)
```
```{r, include = FALSE}
stock_om = readRDS(file.path("../temp", "day_5_1_stock_om_1.RDS"))
```
---

# An EM with an incorrect M assumption

* Create an EM assuming $M = 0.3$
* Fit the EM to data simulated from the OM that assumes $M = 0.2$

```{r, eval = FALSE, results = "hide"}
M_em = list(
  initial_means = array(0.3, dim = c(1,1, length(basic_info$ages)))
)

em_input = prepare_wham_input(
  basic_info = basic_info, 
  selectivity = selectivity_om, 
  NAA_re = NAA_re_om, 
  M = M_em, 
  catch_info = catch_info,
  index_info = index_info,
  F = F_info
)

set.seed(123)
simfit_3 = sim_fn(stock_om, em_input, do.fit = TRUE)
```
```{r, include = FALSE}
simfit_3 = readRDS(file.path("../temp", "day_5_1_simfit_3.RDS"))
```
---

# An EM with an incorrect M assumption

```{r, eval = TRUE, out.width="40%", fig.align="center"}
res <- cbind(simfit_3[[2]]$SSB,simfit_3[[1]]$rep$SSB)
plot(stock_om$years, res[,1], type = 'l', ylab = "SSB", xlab = "Year", ylim = c(0, max(res)))
lines(stock_om$years, res[,2], lty = 2, col = 'red')
```

---

# Closed-loop simulation

* Here we want to
 * use an operating model to simulate the truth
 * use an estimating model to estimate the truth and make catch advice
 * set the operating model to use the catch advice to affect the population

* We need a few functions
 * To determine F to set in the OM from a specified catch
 * To update the OM with the revised F
 * To create the EM input with data from the OM
 * To determine catch advice from the EM
 * To loop through the feedback years of the model (after the base period)

---

# F from Catch

```{r}

get_F_from_catch <- function(om, year, catch, Finit = 0.1, maxF = 10){
    rep = om$rep
    naa = rep$NAA[1,1,year,]
    Maa = rep$MAA[1,1,year,]
    sel_tot = rbind(rep$FAA[,year,]/max(exp(rep$log_FAA_tot[year,]))) #matrix nfleets x n_ages 
    waa = rbind(om$input$data$waa[om$input$data$waa_pointer_fleets, year,]) #matrix nfleets x n_ages 
    nfleets <- length(om$input$data$waa_pointer_fleets)
    get_catch = function(log_F, naa, sel, waa, Maa){
        Faa = exp(log_F) * sel_tot
        Zaa = Maa + apply(Faa,2,sum)
        Catch = 0
        for(a  in 1:length(naa)) for(f in 1:nfleets) Catch = Catch + waa[f,a] * naa[a] * Faa[f,a] *(1 - exp(-Zaa[a]))/Zaa[a];
        return(Catch)
    }
    obj = function(log_F) (catch - get_catch(log_F, naa, sel_tot, waa, Maa))^2
    opt = try(nlminb(log(Finit), obj))
    if(!is.character(opt)) Fsolve = exp(opt$par)[1] else Fsolve = maxF
    if(Fsolve>10) Fsolve = maxF
    print(paste0("Fsolve: ", Fsolve))
    return(Fsolve)
}
```

---

# Update the F in the OM

* Uses `get_F_from_catch`

```{r}
update_om_F = function(om, year, catch){
    rep = om$rep #generate the reported values given the parameters
    year_ind = which(om$years == year) #index corresponding to year
    Fsolve = get_F_from_catch(om, year_ind, catch) #find the F for the catch advice

    #have to be careful if more than one fleet
    FAA = rbind(rep$FAA[,year_ind,]) #n_fleets x n_ages
    FAA_tot <- apply(FAA,2,sum)
    age_ind <- which(FAA_tot == max(FAA_tot))[1]
    selAA = FAA/FAA_tot[age_ind] #sum(sel_all[i,]) = 1
    FAA_catch = Fsolve * selAA
    F_fleet = apply(FAA_catch, 1, max) #full F for each fleet
    if(om$input$data$F_config==1) {
      if(year_ind>1) om$input$par$F_pars[year_ind-1,] <- log(F_fleet) - log(apply(rbind(rep$FAA[,year_ind-1,]),1,max)) #change the F_dev to produce the right full F
      else om$input$par$log_F1[] <- log(F_fleet) #if year is the first year of the model, change F in year 1
    } else{ #alternative configuration of F_pars
      om$input$par$F_pars[year_ind,] <- log(F_fleet)
    }
    om <- fit_wham(om$input, do.fit = FALSE, MakeADFun.silent = TRUE)
    return(om)
}
```

---

# update OM

* Uses `update_om_F`

```{r}
update_om_fn = function(om, seed = 123, interval.info = NULL, random = "log_NAA"){
  obs_names = c("agg_indices","agg_catch","catch_paa","index_paa", "Ecov_obs", "obsvec")

  if(!is.null(interval.info)){ #iteratively update F over assessment interval for the given catch advice
    for(y in interval.info$years){
      om = update_om_F(om, year = y, catch = interval.info$catch) #put in the right F values
      set.seed(seed)
      om_sim = om$simulate(complete=TRUE) #resimulate the population and observations
      om$input$data[obs_names] = om_sim[obs_names] #update any simulated data
      om$input$par[om$input$random] = om_sim[om$input$random] #update any simulated random effects
      # reset the om
      om <- fit_wham(om$input, do.fit = FALSE, MakeADFun.silent = TRUE)
    }
  } else { #otherwise just (re)generate the population
    set.seed(seed)
    om_sim = om$simulate(complete=TRUE) #resimulate the population and observations
    om$input$data[obs_names] = om_sim[obs_names] #update any simulated data
    om$input$par[random] = om_sim[random] #update any simulated random effects
    # reset the om
    om <- fit_wham(om$input, do.fit = FALSE, MakeADFun.silent = TRUE)
  }
  # Loop through feedback period

  return(om)
}
```

---

# Create an EM input

```{r, eval=TRUE}
make_em_input = function(M_em, sel_em, NAA_em, om_data, em_years){
  info = make_info(base_years = em_years) #update the terminal year for the estimation model
  ind_em = 1:length(em_years) #year indices
  #fill in the data from the operating model simulation
  info$catch_info$agg_catch = om_data$agg_catch[ind_em,, drop = FALSE]
  info$index_info$agg_indices = om_data$agg_indices[ind_em,, drop = FALSE]
  info$catch_info$catch_paa = om_data$catch_paa[,ind_em,, drop = FALSE]
  info$index_info$index_paa = om_data$index_paa[,ind_em,, drop = FALSE]
  em_input <- prepare_wham_input(
    basic_info = info$basic_info, 
    selectivity = sel_em, 
    NAA_re = NAA_em, 
    M = M_em, 
    catch_info = info$catch_info,
    index_info = info$index_info)
  return(em_input)
}
```

---

* harvest control rule advises catch at F40
* catch advice is average catch over 5 projection years fishing at F40

```{r, eval=TRUE}
advice_fn = function(em){
  #make 5 year projections using F40. Use average SSB/R and YPR inputs over most recent 5 years
  proj_opts = list(n.yrs=5, use.FXSPR=TRUE, avg.yrs=tail(em$years,5)) 
  em_proj = project_wham(em, do.sdrep = FALSE, proj.opts = proj_opts, MakeADFun.silent=TRUE) #projected version of the em
  advice = mean(apply(em_proj$rep$pred_catch[length(em_proj$years) + 1:5,,drop = FALSE],1,sum)) #mean of the projected catch over the next 5 years fishing at F40
  print(advice)
  return(advice)
}

```

---

# Loop through feedback period

* Uses all of the above functions
* The same seed is used to keep the simulated population and data the same up until F changes

```{r, eval=TRUE}
loop_through_fn = function(om, M_em, selectivity_em, NAA_re_em, assess_years, assess_interval = assess.interval, base_years){
  catches = numeric() #save the catch advice
  for(i in assess_years){
    print(i)
    #make the input for the estimation model

    em_input = make_em_input(M_em, selectivity_em, NAA_re_em, om_data = om$input$data, em_years = base_years[1]:i)
    #fit the estimation model
    em = fit_wham(em_input, do.sdrep = FALSE, do.retro = FALSE, do.osa=FALSE, MakeADFun.silent = TRUE) #no feedback period yet
    #make the catch advice
    advice = advice_fn(em)
    catches = c(catches, advice)
    #set the catch for the next assess_interval years
    interval.info = list(catch = advice, years = i + 1:assess_interval)
    #update the operating model with the right Fs and resimulate the data given those Fs
    om = update_om_fn(om, seed = 123, interval.info = interval.info)  
  }
  return(list(om = om, em  = em, catches = catches))
}
```

---

# Create the OM

* We need to add feedback years to the operating model where F will be replaced based on catch advice
* **Need to remove `random` component of the OM**

```{r, eval = FALSE}
stock_om_info <- make_info(base_years = 1982:2021, n_feedback_years = 40)
basic_info <- stock_om_info$basic_info
catch_info <- stock_om_info$catch_info
index_info <- stock_om_info$index_info
F_info <- stock_om_info$F

stock_om_input = prepare_wham_input(
  basic_info = basic_info, 
  selectivity = selectivity_om, 
  NAA_re = NAA_re_om, 
  M = M_om, 
  catch_info = catch_info,
  index_info = index_info,
  F = F_info
)
stock_om_input$random <- NULL #so inner optimization won't change simulated RE
stock_om_fb <- fit_wham(stock_om_input, do.fit = FALSE, MakeADFun.silent = TRUE)
```
```{r, include = FALSE}
stock_om_fb <- readRDS(file.path("../temp", "day_5_1_stock_om_fb.RDS"))
```

---

# Loop through the feedback period

* assume an assessment every 4 years
* do the closed-loop simulation

```{r, eval=FALSE}

assess.interval = 4
base.years = make_info()$info$years #no feedback period yet
first.year = head(base.years,1)
terminal.year = tail(base.years,1)
assess.years = seq(terminal.year, 
  tail(stock_om_fb$years,1)-assess.interval,
  by = assess.interval)

stock_om_fb = update_om_fn(stock_om_fb)

looped_res = loop_through_fn(stock_om_fb, M_em = M_em, 
  selectivity_em = selectivity_om, 
  NAA_re_em = NAA_re, 
  assess_years = assess.years, 
  base_years = base.years)
looped_rep <- looped_res$om$rep
```
```{r, include = FALSE}
looped_res <- readRDS(file.path("../temp", "day_5_1_looped_res.RDS"))
looped_rep <- looped_res$om$rep
```
---

# Loop through the feedback period

.pull-left[
```{r, eval = TRUE, out.width="80%", fig.align="center"}
res <- cbind(stock_om_fb$rep$SSB, looped_rep$SSB)
plot(stock_om_fb$years, res[,1], type = "l", ylim = c(0,max(res)), ylab = "SSB", xlab = "Year")
lines(stock_om_fb$years, res[,2], col = "red")
```
]
.pull-right[
```{r, eval = TRUE, out.width="80%", fig.align="center"}
res <- cbind(looped_rep$SSB[1:76]/exp(looped_rep$log_SSB_FXSPR[1:76,1]), looped_res$em$rep$SSB[1:76]/exp(looped_res$em$rep$log_SSB_FXSPR[1:76,1]))
plot(stock_om_fb$years[1:76], res[,1], type = "l", ylim = c(0,max(res)), ylab = "SSB/SSB40", xlab = "Year")
lines(stock_om_fb$years[1:76], res[,2], col = "red")
```
]